---
title: "STAT 6302 - Project"
author: Baldwin, Jamie
format: 
  html:
    toc: true
    embed-resources: true
    code-fold: true
    code-line-numbers: true
    code-summary: "SHOW ME THE CODE!"
    code-tools: true
    theme: sketchy
editor: visual
---

```{=html}
<style>
body {
    background-color: #F8C8DC; /* Example background color */
    color: #8B5E83;
    
}

/* Active tab (when selected) */
.nav-tabs .nav-link.active {
    background-color: #d4a2f2e0;  /* Active tab background color */
    color: #8B5E83;               /* Active tab text color */
}

    .highlight {
        background-color: DarkOrchid;
        color: black;
        padding: 2px;
    }

</style>
```

```{r Data Initiation}
autism <- read.csv("C:\\Users\\jasmi\\OneDrive - Southern Methodist University\\Previous Courses\\6302\\Project\\autism+screening+adult\\autism.csv")
head(autism)
```

## Exploratory Data Analysis

### Variable Types & Distributions

```{r EDA 1}
### WE WANT TO CATCH ANY MISSING VALUES AND CORRECTLY CHANGE THEM TO NAs

# Combine "others" into "Others" in the ethnicity variable
autism$ethnicity <- ifelse(autism$ethnicity == "others", "Others", autism$ethnicity)


# Replace "" and "?" with NA across all character or factor columns
autism[] <- lapply(autism, function(x) {
  if (is.character(x) || is.factor(x)) {
    x <- as.character(x)
    x[x == "" | x == "?"] <- NA
    return(as.factor(x))  # Reconvert to factor if desired
  } else {
    return(x)
  }
})

# Also convert numeric "NA-like" values (if stored as strings) to actual NA
autism[] <- lapply(autism, function(x) {
  if (is.character(x)) {
    x[x %in% c("NA", "na", "NaN", "null")] <- NA
    return(as.factor(x))
  } else {
    return(x)
  }
})

### WANT TO RENAME SOME MISSPELLED VARIABLE NAMES

# Rename variables in the dataset
names(autism)[names(autism) == "austim"] <- "autism"
names(autism)[names(autism) == "jundice"] <- "jaundice"
names(autism)[names(autism) == "contry_of_res"] <- "country_of_res"



lapply(autism, unique)

```

```{r EDA 2}
# Load libraries
library(tidyverse)
library(janitor)
library(ggplot2)
library(patchwork)  #For multi-plot layout

#Clean names, just in case
autism <- clean_names(autism)

#Quick structure view
glimpse(autism)

#Create histograms for A1–A10 Scores using a loop
score_vars <- paste0("a", 1:10, "_score")

score_plots <- lapply(score_vars, function(var) {
  ggplot(autism, aes_string(x = var)) +
    geom_histogram(binwidth = 1, fill = "lightsteelblue", color = "black") +
    labs(title = paste0("Distribution of ", toupper(var)), x = var, y = "Count")
})

#Arrange a few sample plots using patchwork
(score_plots[[1]] | score_plots[[2]]) /
(score_plots[[3]] | score_plots[[4]]) /
(score_plots[[5]] | score_plots[[6]])/
(score_plots[[7]] | score_plots[[8]]) /
(score_plots[[9]] | score_plots[[10]])

#Histogram: Age
ggplot(autism, aes(x = age)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
  labs(title = "Age Distribution", x = "Age", y = "Count")

#Histogram: Result Score
ggplot(autism, aes(x = result)) +
  geom_histogram(binwidth = 1, fill = "orchid", color = "black") +
  labs(title = "Total Score Distribution", x = "Result", y = "Count")

#Bar plot: Gender
ggplot(autism, aes(x = gender)) +
  geom_bar(fill = "lightcoral") +
  labs(title = "Gender Distribution", x = "Gender", y = "Count")

#Bar plot: Ethnicity
ggplot(autism, aes(x = ethnicity)) +
  geom_bar(fill = "lightgreen") +
  labs(title = "Ethnicity Distribution", x = "Ethnicity", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Bar plot: Used App Before
ggplot(autism, aes(x = used_app_before)) +
  geom_bar(fill = "lightblue") +
  labs(title = "App Usage History", x = "Used App Before", y = "Count")

ggplot(autism, aes(x = jaundice)) +
  geom_bar(fill = "goldenrod") +
  labs(title = "Distribution of Jaundice at Birth", x = "Jaundice", y = "Count") +
  theme_minimal()

library(forcats)

autism %>%
  count(country_of_res, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = fct_reorder(country_of_res, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Countries of Residence", x = "Country", y = "Count") +
  theme_minimal()

ggplot(autism, aes(x = relation)) +
  geom_bar(fill = "mediumpurple") +
  labs(title = "Survey Respondent Relation", x = "Relation", y = "Count") +
  theme_minimal()

```

**FROM BOTH THE UNIQUE VALUES CODE AND AGE DISTRIBUTION GRAPH ABOVE WE CAN SEE AN INPUTING ERROR IN THE DATA HAS OCCURED WITH AN AGE VALUE OF 383, I WOULD SAY IT IS SAFE TO ASSUME THE CORRECT AGE WAS MEANT TO BE 38. I WILL MAKE THE ASSUMPTION IT WAS MEANT TO BE 38 AS THE REST OF THE AGES RANGE FROM 17 TO 64, AND RECREATE THAT GRAPH.**

```{r EDA 2.5}
#Fixing the typo in age
autism$age[autism$age == 383] <- 38

#Histogram: Age
ggplot(autism, aes(x = age)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
  labs(title = "Age Distribution", x = "Age", y = "Count")

```

### Drop Variables

Since the variable age description (age_desc) only has one level "18 and more" we can remove it completely as it wont add anything to the model.

```{r EDA 4}
# Function to show unique values and their counts
value_counts <- function(var) {
  tbl <- table(var, useNA = "ifany")
  print(tbl)
  cat("\nTotal unique values (including NA):", length(tbl), "\n")
}

# Example usage:
value_counts(autism$age_desc)
value_counts(autism$relation)
value_counts(autism$used_app_before)
value_counts(autism$jaundice)


#Drop the variable 'age_desc'
autism <- autism[, !names(autism) %in% "age_desc"]
autism <- autism[, !names(autism) %in% "used_app_before"]
```

### Missing Data

```{r EDA 3}
#Count of missing values per variable
colSums(is.na(autism))

#Impute age with mean
autism$age[is.na(autism$age)] <- mean(autism$age, na.rm = TRUE)

#Function to compute mode (most frequent value)
get_mode <- function(x) {
  ux <- na.omit(unique(x))
  ux[which.max(tabulate(match(x, ux)))]
}

#Impute ethnicity with mode
autism$ethnicity[is.na(autism$ethnicity)] <- get_mode(autism$ethnicity)

#Impute relation with mode
autism$relation[is.na(autism$relation)] <- get_mode(autism$relation)

```

### Encode

BESURE TO INCLUDE DUMMY INCODING IN RECIPE

### Split into Predictors and Outcome

THIS WILL BE DONE IN THE MODEL RECIPE

### Outcome Variable Balance

```{r EDA 7}
#Convert outcome variable to factor
autism$autism <- as.factor(autism$autism)

#Frequency table
table(autism$autism)

#Proportion table    
prop.table(table(autism$autism))

library(ggplot2)
library(dplyr)

# Bar plot: Count comparison
autism %>%
  select(autism, class_asd) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Response") %>%
  ggplot(aes(x = Response, fill = Variable)) +
  geom_bar(position = "dodge") +
  labs(title = "Comparison of Autism Diagnosis and App Classification",
       x = "Response (Yes/No)", y = "Count") +
  scale_fill_manual(values = c("autism" = "skyblue", "class_asd" = "orchid")) +
  theme_minimal()

```

### Up-Sampling

CAN DO THIS IN RECIPE

### Correlations & Associations

```{r EDA 8}
### numeric variables ###
numeric_vars <- autism %>% select_if(is.numeric)

#Correlation matrix
cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs")

#Plot
library(corrplot)
corrplot::corrplot(cor_matrix, method = "circle", type = "upper", tl.col = "black")

### Categorical Variables ###

# Load required libraries
library(ggplot2)
library(dplyr)
library(patchwork)  # for arranging plots

### Categorical Variables - Chi-square Tests ###
table_gender <- table(autism$gender, autism$autism)
chisq.test(table_gender)

table_ethnicity <- table(autism$ethnicity, autism$autism)
chisq.test(table_ethnicity)

table_relation <- table(autism$relation, autism$autism)
chisq.test(table_relation)

table_jaundice <- table(autism$jaundice, autism$autism)
chisq.test(table_jaundice)

table_cor <- table(autism$country_of_res, autism$autism)
chisq.test(table_cor)

table_class <- table(autism$class_asd, autism$autism)
chisq.test(table_class)

### Stacked Bar Plots ###
p1 <- ggplot(autism, aes(x = gender, fill = autism)) +
  geom_bar(position = "fill") +
  ylab("Proportion") +
  ggtitle("Gender vs Autism") +
  theme_minimal()

p2 <- ggplot(autism, aes(x = ethnicity, fill = autism)) +
  geom_bar(position = "fill") +
  ylab("Proportion") +
  ggtitle("Ethnicity vs Autism") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3 <- ggplot(autism, aes(x = relation, fill = autism)) +
  geom_bar(position = "fill") +
  ylab("Proportion") +
  ggtitle("Relation vs Autism") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p4 <- ggplot(autism, aes(x = jaundice, fill = autism)) +
  geom_bar(position = "fill") +
  ylab("Proportion") +
  ggtitle("Jaundice vs Autism") +
  theme_minimal()

library(forcats)

# Create a new variable with only the top 10 countries
autism <- autism %>%
  mutate(top_country = fct_lump(country_of_res, n = 10))

# Updated stacked bar plot
p5 <- ggplot(autism, aes(x = top_country, fill = autism)) +
  geom_bar(position = "fill") +
  ylab("Proportion") +
  ggtitle("Top 10 Countries vs Autism") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


p6 <- ggplot(autism, aes(x = class_asd, fill = autism)) +
  geom_bar(position = "fill") +
  ylab("Proportion") +
  ggtitle("Classified by App vs Autism") +
  theme_minimal()

### Arrange Plots into Grid ###
(p1 | p3)/(p4 | p6)

(p2 | p5)

```

### Survey Score Patterns

```{r EDA 9}
#Boxplot of result by autism
ggplot(autism, aes(x = autism, y = result, fill = autism)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Total Survey Result by Autism Diagnosis")

#Or use violin plot
ggplot(autism, aes(x = autism, y = result, fill = autism)) +
  geom_violin() +
  theme_minimal() +
  ggtitle("Distribution of Result by Autism Diagnosis")

```

These plots pretty much confirm what we know about scoring for any autism screening survey that it's not perfect and is on more of a sliding scale than a defined one. So for the lack of additional context this section will not be included in the report.

### Interaction Exploration

```{r EDA 10}
# Load required library
library(dplyr)

# Set significance threshold
alpha <- 0.05

# Get predictor names (exclude only the outcome variable)
predictors <- setdiff(names(autism), c("autism", "top_country"))

# Initialize result dataframe
significant_interactions <- data.frame(var1 = character(), var2 = character(), p_value = numeric())

# Loop over all unique pairs of predictors
for (i in 1:(length(predictors) - 1)) {
  for (j in (i + 1):length(predictors)) {
    var1 <- predictors[i]
    var2 <- predictors[j]
    
    # Try-catch to handle problematic combinations
    try({
      # Main effects only model
      main_formula <- as.formula(paste("autism ~", var1, "+", var2))
      model_main <- glm(main_formula, data = autism, family = binomial)

      # Interaction model
      interaction_formula <- as.formula(paste("autism ~", var1, "*", var2))
      model_inter <- glm(interaction_formula, data = autism, family = binomial)

      # Likelihood ratio test
      lrt <- anova(model_main, model_inter, test = "LRT")
      p_val <- lrt$`Pr(>Chi)`[2]

      if (!is.na(p_val) && p_val < alpha) {
        significant_interactions <- rbind(significant_interactions, 
                                          data.frame(var1 = var1, var2 = var2, p_value = p_val))
      }
    }, silent = TRUE)
  }
}

# Sort by p-value and print results
significant_interactions <- significant_interactions %>% arrange(p_value)
print(significant_interactions)


```

```{r EDA 11}
library(ggplot2)
library(dplyr)
library(patchwork)

# Your significant interactions
interactions <- data.frame(
  var1 = c("a6_score", "gender", "a8_score", "a4_score", "a10_score", "a8_score", "ethnicity", "a8_score"),
  var2 = c("ethnicity", "jaundice", "a9_score", "a8_score", "jaundice", "gender", "result", "country_of_res")
)

# Function to generate interaction plots and return ggplot object
generate_plot <- function(data, var1, var2) {
  v1 <- data[[var1]]
  v2 <- data[[var2]]
  
  is_v1_cat <- is.factor(v1) || is.character(v1)
  is_v2_cat <- is.factor(v2) || is.character(v2)

  if (!is_v1_cat & !is_v2_cat) {
    return(
      ggplot(data, aes_string(x = var1, y = var2, color = "autism")) +
        geom_point(alpha = 0.6, position = position_jitter(width = 0.2, height = 0.2)) +
        theme_minimal() +
        labs(title = paste(var1, "vs", var2))
    )
  } else if (is_v1_cat & is_v2_cat) {
    return(
      ggplot(data, aes_string(x = var1, fill = "autism")) +
        geom_bar(position = "fill") +
        facet_wrap(as.formula(paste("~", var2))) +
        scale_y_continuous(labels = scales::percent_format()) +
        theme_minimal() +
        labs(y = "Proportion", title = paste(var1, "by", var2))
    )
  } else {
    if (is_v1_cat) {
      return(
        ggplot(data, aes_string(x = var1, y = var2, fill = "autism")) +
          geom_boxplot() +
          theme_minimal() +
          labs(title = paste(var1, "vs", var2))
      )
    } else {
      return(
        ggplot(data, aes_string(x = var2, y = var1, fill = "autism")) +
          geom_boxplot() +
          theme_minimal() +
          labs(title = paste(var2, "vs", var1))
      )
    }
  }
}

# Generate all plots and store them
plots <- list()
for (i in 1:nrow(interactions)) {
  plots[[i]] <- generate_plot(autism, interactions$var1[i], interactions$var2[i])
}

# Display in grid: 2 columns
wrap_plots(plots, ncol = 2)

```

## Modeling

### Base Logistic Regression Model

```{r Model 1}
library(recipes)
library(rsample)
library(parsnip)
library(workflows)
library(tune)
library(yardstick)
library(broom)
##The outcome is a factor
class(autism$autism)

##We will be predicting the P(yes)
levels(autism$autism)

autism_recipe <- recipe(autism ~ a1_score + a2_score + a3_score + a4_score + a5_score + a6_score + a7_score + a8_score + a9_score + a10_score + age + gender + ethnicity + jaundice + country_of_res + relation + class_asd, 
                data = autism) %>% 
  step_dummy(all_nominal_predictors())

set.seed(123)
autism_kfolds <- vfold_cv(autism, v = 10, repeats = 5, strata = "autism")

autism_logistic <- logistic_reg() %>% 
                  set_mode("classification") %>%
                  set_engine("glm") 

autism_wf <- workflow() %>% 
            add_recipe(autism_recipe) %>% 
            add_model(autism_logistic)

set.seed(123)
autism_resammple <- fit_resamples(
              autism_wf,
              resamples = autism_kfolds,
              metrics = metric_set(accuracy, roc_auc, kap, sens, spec, brier_class))

autism_resammple %>% collect_metrics()

##Extract coefficients (after fitting on full data)
autism_model <- autism_wf %>%
  fit(autism) %>%
  extract_fit_parsnip() %>%
  tidy()

autism_model
```

### Custom Model 1

```{r Model 2}
library(themis) # for upsample
autism_recipe2 <- recipe(autism ~ a1_score + a2_score + a3_score + a4_score + a5_score + a6_score + a7_score + a8_score + a9_score + a10_score + age + gender + ethnicity + jaundice + country_of_res + relation + class_asd + result, 
                data = autism) %>% 
  step_interact(terms = ~ ethnicity:result) %>%
  step_interact(terms = ~ gender:jaundice) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_upsample(autism, over_ratio = 1)

set.seed(123)
autism_kfolds <- vfold_cv(autism, v = 10, repeats = 5, strata = "autism")

autism_logistic <- logistic_reg() %>% 
                  set_mode("classification") %>%
                  set_engine("glm") 

autism_wf2 <- workflow() %>% 
            add_recipe(autism_recipe2) %>% 
            add_model(autism_logistic)

set.seed(123)
autism_resammple2 <- fit_resamples(
              autism_wf2,
              resamples = autism_kfolds,
              metrics = metric_set(accuracy, roc_auc, kap, sens, spec, brier_class))

autism_resammple2 %>% collect_metrics()

##Extract coefficients (after fitting on full data)
autism_model2 <- autism_wf2 %>%
  fit(autism) %>%
  extract_fit_parsnip() %>%
  tidy()

autism_model2
```

Adds upsampling and interaction terms

### Custom Model 2

Adds elastic net

```{r Model 3}
library(themis) # for upsample
library(dials)

# Define recipe with interactions and upsampling
autism_recipe3 <- recipe(autism ~ a1_score + a2_score + a3_score + a4_score + a5_score + a6_score + a7_score + a8_score + a9_score + a10_score + age + gender + ethnicity + jaundice + country_of_res + relation + class_asd + result, 
                data = autism) %>% 
  step_interact(terms = ~ ethnicity:result + gender:jaundice) %>%
  step_dummy(all_nominal_predictors()) %>% 
  step_upsample(autism, over_ratio = 1)

# Cross-validation setup
set.seed(123)
autism_kfolds <- vfold_cv(autism, v = 10, repeats = 5, strata = "autism")

# Define penalized logistic regression model with tunable hyperparameters
autism_en_model <- logistic_reg(penalty = tune(), mixture = tune()) %>% 
  set_mode("classification") %>%
  set_engine("glmnet") 

# Define workflow
autism_en_wflow <- workflow() %>% 
  add_recipe(autism_recipe3) %>% 
  add_model(autism_en_model)

# Create tuning grid (space-filling design)
en_param <- extract_parameter_set_dials(autism_en_model)
glh_en <- grid_space_filling(en_param, size = 10, type = "latin_hypercube")

# Tune the model with resampling
set.seed(123)
autism_en_tune <- tune_grid(
  autism_en_wflow,
  resamples = autism_kfolds,
  grid = glh_en,
  metrics = metric_set(accuracy, roc_auc, kap, sens, spec, brier_class)
)

# Collect and print metrics from tuning
autism_en_tune %>% collect_metrics()

# Select best parameters based on ROC AUC
autism_en_param <- autism_en_tune %>% select_best(metric = "roc_auc")

# Finalize the model with best parameters
autism_en_model2 <- finalize_model(autism_en_model, autism_en_param)

# Update the workflow with the finalized model
autism_en_wflow2 <- autism_en_wflow %>%
  update_model(autism_en_model2)

# Fit on resamples to evaluate metrics
autism_resample_final <- fit_resamples(
  autism_en_wflow2,
  resamples = autism_kfolds,
  metrics = metric_set(accuracy, roc_auc, kap, sens, spec, brier_class)
)

# Collect metrics after using the best parameters
autism_resample_final %>%
  collect_metrics()

# Fit the finalized workflow to the full data
autism_en_final_fit <- autism_en_wflow2 %>%
  fit(data = autism)

# Collect and tidy the coefficients
autism_en_params <- autism_en_final_fit %>%
  extract_fit_parsnip() %>%
  tidy()

# Print the coefficients
autism_en_params

library(dplyr)

# Filter out predictors with zero coefficients
autism_en_params_nonzero <- autism_en_params %>%
  filter(estimate != 0)

# Print the filtered results
autism_en_params_nonzero
```

### Elastic Net

```{r Modeling 4}
library(themis) # for upsample

# Define recipe with interactions and upsampling
autism_recipe4 <- recipe(autism ~ a1_score + a2_score + a3_score + a4_score + a5_score + a6_score + a7_score + a8_score + a9_score + a10_score + age + gender + ethnicity + jaundice + country_of_res + relation + class_asd, 
                data = autism) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_upsample(autism, over_ratio = 1)

# Cross-validation setup
set.seed(123)
autism_kfolds <- vfold_cv(autism, v = 10, repeats = 5, strata = "autism")

# Define penalized logistic regression model with tunable hyperparameters
autism_en_model <- logistic_reg(penalty = tune(), mixture = tune()) %>% 
  set_mode("classification") %>%
  set_engine("glmnet") 

# Define workflow
autism_en_wflow3 <- workflow() %>% 
  add_recipe(autism_recipe4) %>% 
  add_model(autism_en_model)

# Create tuning grid (space-filling design)
en_param <- extract_parameter_set_dials(autism_en_model)
glh_en <- grid_space_filling(en_param, size = 10, type = "latin_hypercube")

# Tune the model with resampling
set.seed(123)
autism_en_tune <- tune_grid(
  autism_en_wflow3,
  resamples = autism_kfolds,
  grid = glh_en,
  metrics = metric_set(accuracy, roc_auc, kap, sens, spec, brier_class)
)

# Collect and print metrics from tuning
autism_en_tune %>% collect_metrics()

# Select best parameters based on ROC AUC
autism_en_param3 <- autism_en_tune %>% select_best(metric = "roc_auc")

# Finalize the model with best parameters
autism_en_model3 <- finalize_model(autism_en_model, autism_en_param3)

# Update the workflow with the finalized model
autism_en_wflow4 <- autism_en_wflow3 %>%
  update_model(autism_en_model3)

# Fit on resamples to evaluate metrics
autism_resample_final3 <- fit_resamples(
  autism_en_wflow4,
  resamples = autism_kfolds,
  metrics = metric_set(accuracy, roc_auc, kap, sens, spec, brier_class)
)

# Collect metrics after using the best parameters
autism_resample_final3 %>%
  collect_metrics()

# Fit the finalized workflow to the full data
autism_en_final_fit3 <- autism_en_wflow4 %>%
  fit(data = autism)

# Collect and tidy the coefficients
autism_en_params3 <- autism_en_final_fit3 %>%
  extract_fit_parsnip() %>%
  tidy()

# Print the coefficients
autism_en_params3

library(dplyr)

# Filter out predictors with zero coefficients
autism_en_params_nonzero3 <- autism_en_params3 %>%
  filter(estimate != 0)

# Print the filtered results
autism_en_params_nonzero3
```

No interaction terms

## Additional Code for Report Graphics

```{r Fancy Table}
library(gt)

# Add a column to identify each model
autism_metrics <- autism_resammple %>% collect_metrics() %>% mutate(Model = "Base Model")
autism_metrics2 <- autism_resammple2 %>% collect_metrics() %>% mutate(Model = "Custom Model 1")
autism_metrics3 <- autism_resample_final %>% collect_metrics() %>% mutate(Model = "Custom Model 2")
autism_metrics4 <- autism_resample_final3 %>% collect_metrics() %>% mutate(Model = "EN Model")

# Combine all into one data frame
all_metrics <- bind_rows(autism_metrics, autism_metrics2, autism_metrics3, autism_metrics4)

# Pivot wider to compare metrics across models
metrics_wide <- all_metrics %>%
  select(Model, .metric, mean) %>%
  pivot_wider(names_from = .metric, values_from = mean)

# Print the final comparison table using gt
metrics_wide %>%
  gt() %>%
  tab_header(
    title = "Model Comparison: Performance Metrics",
    subtitle = "Evaluating classification models on Autism Screening"
  ) %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 3
  ) %>%
  cols_label(
    Model = "Model",
    accuracy = "Accuracy",
    kap = "Kappa",
    roc_auc = "ROC AUC",
    sens = "Specificity",
    spec = "Sensitivity",
    brier_class = "Brier Score"
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  opt_table_outline() %>%
  opt_row_striping()

```

### Base Model

```{r Base Model Powerful Predictors}
library(dplyr)
library(ggplot2)


# Get top 5 positive and top 5 negative predictors
top_positivea <- autism_model %>%
  arrange(desc(estimate)) %>%
  slice_head(n = 5)

top_negativea <- autism_model %>%
  arrange(estimate) %>%
  slice_head(n = 5)

# Combine and plot
top_combineda <- bind_rows(top_positivea, top_negativea)

ggplot(top_combineda, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 Positive and Negative Predictors",
    x = "Predictor",
    y = "Coefficient Estimate"
  ) +
  scale_fill_manual(values = c("TRUE" = "#66c2a5", "FALSE" = "#fc8d62")) +
  theme_minimal()

```

```{r Base Model Predictors List}
library(dplyr)

# Count and list positive and negative predictors
positive_predictorsa <- autism_model %>%
  filter(term != "(Intercept)", estimate > 0) %>%
  arrange(desc(estimate))

negative_predictorsa <- autism_model %>%
  filter(term != "(Intercept)", estimate < 0) %>%
  arrange(estimate)

# Print counts
cat("Number of positive predictors:", nrow(positive_predictorsa), "\n")
cat("Number of negative predictors:", nrow(negative_predictorsa), "\n\n")

# List predictors
cat("Positive predictors:\n")
print(positive_predictorsa)

cat("\nNegative predictors:\n")
print(negative_predictorsa)

```

### Custom Model 1

```{r Custom Model 1 Powerful Predictors}
library(dplyr)
library(ggplot2)


# Get top 5 positive and top 5 negative predictors
top_positiveb <- autism_model2 %>%
  arrange(desc(estimate)) %>%
  slice_head(n = 5)

top_negativeb <- autism_model2 %>%
  arrange(estimate) %>%
  slice_head(n = 5)

# Combine and plot
top_combinedb <- bind_rows(top_positiveb, top_negativeb)

ggplot(top_combinedb, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 Positive and Negative Predictors",
    x = "Predictor",
    y = "Coefficient Estimate"
  ) +
  scale_fill_manual(values = c("TRUE" = "#66c2a5", "FALSE" = "#fc8d62")) +
  theme_minimal()

```

```{r Custom Model 1 Predictors List}
library(dplyr)

#Count and list positive and negative predictors
positive_predictorsb <- autism_model2 %>%
  filter(term != "(Intercept)", estimate > 0) %>%
  arrange(desc(estimate))

negative_predictorsb <- autism_model2 %>%
  filter(term != "(Intercept)", estimate < 0) %>%
  arrange(estimate)

#Print counts
cat("Number of positive predictors:", nrow(positive_predictorsb), "\n")
cat("Number of negative predictors:", nrow(negative_predictorsb), "\n\n")

# List predictors
cat("Positive predictors:\n")
print(positive_predictorsb)

cat("\nNegative predictors:\n")
print(negative_predictorsb)

```

### Custom Model 2

```{r Custom Model 2 Powerful Predictors}
library(dplyr)
library(ggplot2)


# Get top 5 positive and top 5 negative predictors
top_positivec <- autism_en_params_nonzero %>%
  arrange(desc(estimate)) %>%
  slice_head(n = 5)

top_negativec <- autism_en_params_nonzero %>%
  filter(term != "(Intercept)") %>%
  arrange(estimate) %>%
  slice_head(n = 5)

# Combine and plot
top_combinedc <- bind_rows(top_positivec, top_negativec)

ggplot(top_combinedc, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 Positive and Negative Predictors",
    x = "Predictor",
    y = "Coefficient Estimate"
  ) +
  scale_fill_manual(values = c("TRUE" = "#66c2a5", "FALSE" = "#fc8d62")) +
  theme_minimal()

```

```{r Custom Model 2 Predictors List}
library(dplyr)

#Count and list positive and negative predictors
positive_predictorsc <- autism_en_params_nonzero %>%
  filter(term != "(Intercept)", estimate > 0) %>%
  arrange(desc(estimate))

negative_predictorsc <- autism_en_params_nonzero %>%
  filter(term != "(Intercept)", estimate < 0) %>%
  arrange(estimate)

#Print counts
cat("Number of positive predictors:", nrow(positive_predictorsc), "\n")
cat("Number of negative predictors:", nrow(negative_predictorsc), "\n\n")

# List predictors
cat("Positive predictors:\n")
print(positive_predictorsc)

cat("\nNegative predictors:\n")
print(negative_predictorsc)

```

### Elastic Net

```{r EN Powerful Predictors}
library(dplyr)
library(ggplot2)


# Get top 5 positive and top 5 negative predictors
top_positived <- autism_en_params_nonzero3 %>%
  arrange(desc(estimate)) %>%
  slice_head(n = 5)

top_negatived <- autism_en_params_nonzero3 %>%
  filter(term != "(Intercept)") %>%
  arrange(estimate) %>%
  slice_head(n = 5)

# Combine and plot
top_combinedd <- bind_rows(top_positived, top_negatived)

ggplot(top_combinedd, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 5 Positive and Negative Predictors",
    x = "Predictor",
    y = "Coefficient Estimate"
  ) +
  scale_fill_manual(values = c("TRUE" = "#66c2a5", "FALSE" = "#fc8d62")) +
  theme_minimal()

```

```{r EN Predictors List}
library(dplyr)

#Count and list positive and negative predictors
positive_predictorsd <- autism_en_params_nonzero3 %>%
  filter(term != "(Intercept)", estimate > 0) %>%
  arrange(desc(estimate))

negative_predictorsd <- autism_en_params_nonzero3 %>%
  filter(term != "(Intercept)", estimate < 0) %>%
  arrange(estimate)

#Print counts
cat("Number of positive predictors:", nrow(positive_predictorsd), "\n")
cat("Number of negative predictors:", nrow(negative_predictorsd), "\n\n")

# List predictors
cat("Positive predictors:\n")
print(positive_predictorsd)

cat("\nNegative predictors:\n")
print(negative_predictorsd)

```

## Edits (9/24/25)

```{r Regression Summary Table to Graph}
# Load libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Create your data frame
df <- data.frame(
  Model = c("Base Model", "Custom Model 1", "Custom Model 2", "EN Model"),
  Accuracy = c(0.85, 0.70, 0.68, 0.69),
  Brier_Score = c(0.12, 0.22, 0.20, 0.20),
  Kappa = c(0.12, 0.16, 0.18, 0.19),
  ROC_AUC = c(0.62, 0.61, 0.71, 0.71),
  Specificity = c(0.95, 0.72, 0.69, 0.69),
  Sensitivity = c(0.14, 0.53, 0.63, 0.63)
)

# Reshape to long format for ggplot
df_long <- df %>%
  pivot_longer(cols = -Model, names_to = "Metric", values_to = "Value")

# Plot: small multiples (facet by metric)
ggplot(df_long, aes(x = Model, y = Value, fill = Model)) +
  geom_col(position = "dodge") +
  facet_wrap(~ Metric, scales = "free_y") +
  geom_text(aes(label = round(Value, 3)), 
            position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  labs(title = "Model Comparison: Performance Metrics",
       subtitle = "Autism Screening Classification Models",
       x = "", y = "Value") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "none")

```

